Fractional Types

### 2016-12-19 typespp.tex
Introduce notation for ∀∃

\usepackage{amsthm,amstext,amssymb,amsmath}
\usepackage{datetime}
\usepackage{fdsymbol} %% must be loaded before bbold because it interferes with numbers
\usepackage{bbold}
\usepackage{url}
\usepackage{xcolor}
\usepackage{multicol}
\usepackage{proof}
\usepackage{stmaryrd}
\usepackage{bussproofs}
\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{quotes}
\usepackage[section]{placeins}
\usepackage{braket}

@@

@@

\unitetl/ :& \mbox{\AE}{\overline{k}}.[\order{\idiso} \circledast \mathbb{T},

@@

                   \forall\exists{\overline{k}}.[\mathbb{T},v]

@@

\unitetr/ :& \forall\exists{\overline{k}}.[\mathbb{T} \circledast \order{\idiso},

@@

                   \forall\exists{\overline{k}}.[\mathbb{T},v]

@@

\swapt/ :& \forall\exists{\overline{k}}.[\mathbb{T}_1 \circledast \mathbb{T}_2,

@@

                  \forall\exists{\overline{k}}.[\mathbb{T}_2 \circledast \mathbb{T}_1,

@@

  \assoclt/ :& \forall\exists{\overline{k}}.

@@

                      \forall\exists{\overline{k}}.

@@

  \forall\exists{k}.[\iorder{c} \circledast \order{c}, \langle 1/c^k, c^k  \rangle]

@@

  \forall\exists{k}.[\order{c} \circledast \iorder{c}, \langle c^k,1/c^k \rangle]

@@

\synchl :& \forall\exists{k}.[(\order{c}\circledast\iorder{c})\circledast\order{c},

@@

\synchr :& \forall\exists{k}.[\order{c}\circledast(\iorder{c}\circledast\order{c}),

@@

{\jdg{}{}{\idiso/ : \forall\exists{\overline{k}}.[\mathbb{T},v] \leftrightblackspoon
   \forall\exists{\overline{k}}.[\mathbb{T},v]}}

@@

  \forall\exists{\overline{k_1}}.[\mathbb{T}_1,v_1] \leftrightblackspoon
  \forall\exists{\overline{k_2}}.[\mathbb{T}_2,v_2]} \quad

@@

  \forall\exists{\overline{k_2}}.[\mathbb{T}_2,v_2] \leftrightblackspoon
  \forall\exists{\overline{k_3}}.[\mathbb{T}_3,v_3]}

@@

 \forall\exists{\overline{k_1}}.[\mathbb{T}_1,v_1] \leftrightblackspoon
 \forall\exists{\overline{k_3}}.[\mathbb{T}_3},v_3]}

@@

 \forall\exists{\overline{k_1}}.[\mathbb{T}_1,v_1] \leftrightblackspoon
 \forall\exists{\overline{k_2}}.[\mathbb{T}_2,v_2]}

@@

 \forall\exists{\overline{k_3}}.[\mathbb{T}_3,v_3] \leftrightblackspoon
 \forall\exists{\overline{k_4}}.[\mathbb{T}_4,v_4]}

@@

  \forall\exists{\overline{k_1k_3}}.[\mathbb{T}_1 \circledast \mathbb{T}_3, \langle v_1,v_3 \rangle]

@@

  \forall\exists{\overline{k_2k_4}}.[\mathbb{T}_2 \circledast \mathbb{T}_4}, \langle v_2,v_4 \rangle]}

@@

\forall\exists{k}.\mathbb{T_2}_\bullet$ introduces a $\forall$ in the left-to-right

@@

$\forall\exists{k}.\mathbb{T_1}_\bullet \leftrightblackspoon

@@

$\forall\exists{k}.\mathbb{T_1}_\bullet \leftrightblackspoon
\forall\exists{k}.\mathbb{T_2}_\bullet$ maintains (think of elimination then

@@

In the new definition, we enrich the types to pointed types $\mathbb{T_\bullet}$
where each type $\mathbb{T}$ refers to a particular value $v$ in
$\mathbb{T}$. As all fractional types are inhabited (see Remark in
Sec.~\ref{sec:symmetryg}), there is no loss in generality in going to pointed
types. We then allow universal quantification over indices $k$ associated to
values, over a collection of types.  In particular, the type
$\forall k.[\order{c} \circledast \iorder{c}, \langle c^k, 1/c^k \rangle]$
imposes a constraint on the two values associated with the types $\order{c}$ and
$\iorder{c}$; they must be ``synchronized'' by sharing the same~$k$ but are
otherwise arbitrary. Once we introduce a universal quantifier in one direction
of execution, we are led to also include an existential quantifier for the
reverse execution. Intuitively, a forward evaluation step that is at liberty to
choose a $k$ will, when reversed encounter a particular, i.e., existentially
quantified, $k$. A type $\mathbb{T}_\bullet$ is therefore generally of the form
$\forall\exists{\overline{k}}.[\mathbb{T},v]$ where all the quantifiers are
lifted to the top followed by a non-dependent type and a value in that type.

The adaptation of the combinators to the pointed dependent setting in Fig .~\ref
{pifracd:comb} is relatively straightforward. First, the types of the
combinators must keep track of how the value currently in focus changes during
evaluation, and hence directly encode the operational semantics of the
language. As suggested above, the quantifiers are interpreted differently
depending on which direction the rule is used. We use a $\forall\exists$
notation to express the presence of \emph{one} quantifier that is interpreted
differently in each direction. The notation has the advantage of being compact,
avoiding duplication, and making the symmetry of the rules explicit. We also
include two new pairs of combinators $\synchl/\packl$ and $\synchr/\packr$ that
serve as explicit quantifier introduction and elimination rules: they serve as
explicit ``synchronization'' points as illustrated in the example below.

>>>>

\usepackage{amsthm,amstext,amssymb,amsmath}
\usepackage{datetime}
\usepackage{fdsymbol} %% must be loaded before bbold because it interferes with numbers
\usepackage{bbold}
\usepackage{url}
\usepackage{xcolor}
\usepackage{multicol}
\usepackage{proof}
\usepackage{stmaryrd}
\usepackage{bussproofs}
\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{quotes}
\usepackage[section]{placeins}
\usepackage{braket}
\usepackage{mathtools}

@@

\newcommand{\quant}[0]{\mbox{\ensuremath{\mathrlap{\diamond}\square}}}

@@

\unitetl/ :& \quant{\overline{k}}.[\order{\idiso} \circledast \mathbb{T},

@@

                   {\overline{k}}.[\mathbb{T},v]

@@

\unitetr/ :& \quant{\overline{k}}.[\mathbb{T} \circledast \order{\idiso},

@@

                   \quant{\overline{k}}.[\mathbb{T},v]

@@

\swapt/ :& \quant{\overline{k}}.[\mathbb{T}_1 \circledast \mathbb{T}_2,

@@

                  \quant{\overline{k}}.[\mathbb{T}_2 \circledast \mathbb{T}_1,

@@

  \assoclt/ :& \quant{\overline{k}}.

@@

                      \quant{\overline{k}}.

@@

  \quant{k}.[\iorder{c} \circledast \order{c}, \langle 1/c^k, c^k  \rangle]

@@

  \quant{k}.[\order{c} \circledast \iorder{c}, \langle c^k,1/c^k \rangle]

@@

\synchl :& \quant{k}.[(\order{c}\circledast\iorder{c})\circledast\order{c},

@@

\synchr :& \quant{k}.[\order{c}\circledast(\iorder{c}\circledast\order{c}),

@@

{\jdg{}{}{\idiso/ : \quant{\overline{k}}.[\mathbb{T},v] \leftrightblackspoon
   \quant{\overline{k}}.[\mathbb{T},v]}}

@@

  \quant{\overline{k_1}}.[\mathbb{T}_1,v_1] \leftrightblackspoon
  \quant{\overline{k_2}}.[\mathbb{T}_2,v_2]} \quad

@@

  \quant{\overline{k_2}}.[\mathbb{T}_2,v_2] \leftrightblackspoon
  \quant{\overline{k_3}}.[\mathbb{T}_3,v_3]}

@@

 \quant{\overline{k_1}}.[\mathbb{T}_1,v_1] \leftrightblackspoon
 \quant{\overline{k_3}}.[\mathbb{T}_3},v_3]}

@@

 \quant{\overline{k_1}}.[\mathbb{T}_1,v_1] \leftrightblackspoon
 \quant{\overline{k_2}}.[\mathbb{T}_2,v_2]}

@@

 \quant{\overline{k_3}}.[\mathbb{T}_3,v_3] \leftrightblackspoon
 \quant{\overline{k_4}}.[\mathbb{T}_4,v_4]}

@@

  \quant{\overline{k_1k_3}}.[\mathbb{T}_1 \circledast \mathbb{T}_3, \langle v_1,v_3 \rangle]

@@

  \quant{\overline{k_2k_4}}.[\mathbb{T}_2 \circledast \mathbb{T}_4}, \langle v_2,v_4 \rangle]}

@@

\quant{k}.\mathbb{T_2}_\bullet$ introduces a $\forall$ in the left-to-right

@@

$\quant{k}.\mathbb{T_1}_\bullet \leftrightblackspoon

@@

$\quant{k}.\mathbb{T_1}_\bullet \leftrightblackspoon
\quant{k}.\mathbb{T_2}_\bullet$ maintains (think of elimination then

@@

In the new definition, we enrich the types to pointed types $\mathbb{T_\bullet}$
where each type $\mathbb{T}$ refers to a particular value $v$ in
$\mathbb{T}$. As all fractional types are inhabited (see Remark in
Sec.~\ref{sec:symmetryg}), there is no loss in generality in going to pointed
types. We then allow universal quantification over indices $k$ associated to
values, over a collection of types.  In particular, the type
$\forall k.[\order{c} \circledast \iorder{c}, \langle c^k, 1/c^k \rangle]$
imposes a constraint on the two values associated with the types $\order{c}$ and
$\iorder{c}$; they must be ``synchronized'' by sharing the same~$k$ but are
otherwise arbitrary. Once we introduce a universal quantifier in one direction
of execution, we are led to also include an existential quantifier for the
reverse execution. Intuitively, a forward evaluation step that is at liberty to
choose a $k$ will, when reversed encounter a particular, i.e., existentially
quantified, $k$. A type $\mathbb{T}_\bullet$ is therefore generally of the form
$\quant{\overline{k}}.[\mathbb{T},v]$ where all the quantifiers are
lifted to the top followed by a non-dependent type and a value in that type.

The adaptation of the combinators to the pointed dependent setting in Fig .~\ref
{pifracd:comb} is relatively straightforward. First, the types of the
combinators must keep track of how the value currently in focus changes during
evaluation, and hence directly encode the operational semantics of the
language. As suggested above, the quantifiers are interpreted differently
depending on which direction the rule is used. We use a $\quant$
notation to express the presence of \emph{one} quantifier that is interpreted
differently in each direction. The notation has the advantage of being compact,
avoiding duplication, and making the symmetry of the rules explicit. We also
include two new pairs of combinators $\synchl/\packl$ and $\synchr/\packr$ that
serve as explicit quantifier introduction and elimination rules: they serve as
explicit ``synchronization'' points as illustrated in the example below.
